Boosting Higgs Pair Production in the $b\bar{b}b\bar{b}$ Channel at the LHC with Multivariate Techniques

### 2015-11-17 paper/sec-mva.tex
more work on paper
Even more work on paper

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
      \vspace{-1cm}
  \includegraphics[width=0.90\textwidth]{plots/res_nnarch.pdf}
  \vspace{-1cm}
\caption{\small Architecture of the ANN used for the analysis of the resolved
  category, with $N_{\rm var}=13$ input variables and thus the same number of neurons
  in the first layer.
  %
  The color code in the neuron connections (the weights) is a heat map,
  with red indicating large values and black small values.
}
\label{fig:nnarch}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{Signal significance after the MVA}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
\includegraphics[width=0.65\textwidth]{plots/Boosted_disc.pdf}
\includegraphics[width=0.48\textwidth]{plots/Intermediate_disc.pdf}
\includegraphics[width=0.48\textwidth]{plots/Resolved_disc.pdf}
\caption{\small The distribution as a function of the ANN output
  for the signal and background MC events in the three categories:
  boosted (upper plot), intermediate (lower left plot) and
  resolved (lower right plot).
  %
  All distributions are normalized so that their integral
  adds up to one.
}
\label{fig:nnresponse}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
\includegraphics[width=0.65\textwidth]{plots/nev2.pdf}
\caption{\small Number of signal (dashed lines) and background (solid lines)
  events expected at the HL-LHC as a function of the cut in the ANN output,
  for the three separate categories.
}
\label{fig:nev2}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
\includegraphics[width=0.49\textwidth]{plots/res_wgthist.pdf}
\includegraphics[width=0.49\textwidth]{plots/boosted_wgthist.pdf}
\vspace{-0.5cm}
\caption{\small
Distribution of the total associated weight,
Eq.~(\ref{eq:totweight}) for each of the $N_{\rm var}$ input
variables of the resolved (left plot) and boosted (right plot) categories.
%
See Sect.~\ref{sec:input} for a description of each of the input
variables to the MVA.
}
\label{fig:nnweights}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
\includegraphics[width=0.48\textwidth]{plots/ssb.pdf}
\includegraphics[width=0.48\textwidth]{plots/sb.pdf}
\caption{\small
  The values of the signal significance $S/\sqrt{B}$ and of the
  signal over background ratio $S/B$ for the boosted, intermediate
  and resolved categories as  function of the cut
  $y_{\rm cut}=0$ in the ANN output.
  %
  The $y_{\rm cut}=0$
  values correspond to the  cut-based analysis
  results from Table~\ref{table:cutflow}.
}
\label{fig:sb_mva}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
  \begin{center}
      \vspace{-1cm}
  \includegraphics[width=0.90\textwidth]{plots/bst_nnarch_noPU.pdf}
  \vspace{-1cm}
  \caption{\small Architecture of the ANN used for the analysis of the
    boosted
    category, with $N_{\rm var}=21$ input variables and thus
    the same number of neurons
  in the first layer.
  %
  The color code in the neuron connections (the weights) is a heat map,
  with red indicating large values and black small values.
}
\label{fig:nnarch}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

\subsection{Signal significance after the MVA}
\label{sec:signalsignificance}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
\includegraphics[width=0.65\textwidth]{plots/Boosted_disc_noPU.pdf}
\includegraphics[width=0.48\textwidth]{plots/Intermediate_disc_noPU.pdf}
\includegraphics[width=0.48\textwidth]{plots/Resolved_disc_noPU.pdf}
\caption{\small The distribution as a function of the ANN output
  for the signal and background MC events in the three categories:
  boosted (upper plot), intermediate (lower left plot) and
  resolved (lower right plot).
  %
  All distributions are normalized so that their integral
  adds up to one.
}
\label{fig:nnresponse}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
\includegraphics[width=0.65\textwidth]{plots/nev2_noPU.pdf}
\caption{\small Number of signal (dashed lines) and background (solid lines)
  events expected at the HL-LHC as a function of the cut in the ANN output,
  for the three separate categories.
}
\label{fig:nev2}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
\includegraphics[width=0.49\textwidth]{plots/res_wgthist_noPU.pdf}
\includegraphics[width=0.49\textwidth]{plots/bst_wgthist_noPU.pdf}
\vspace{-0.5cm}
\caption{\small
Distribution of the total associated weight,
Eq.~(\ref{eq:totweight}) for each of the $N_{\rm var}$ input
variables of the resolved (left plot) and boosted (right plot) categories.
%
See Sect.~\ref{sec:input} for a description of each of the input
variables to the MVA.
}
\label{fig:nnweights}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

@@

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
\includegraphics[width=0.48\textwidth]{plots/ssb_noPU.pdf}
\includegraphics[width=0.48\textwidth]{plots/sb_noPU.pdf}
\caption{\small
  The values of the signal significance $S/\sqrt{B}$ and of the
  signal over background ratio $S/B$ for the boosted, intermediate
  and resolved categories as  function of the cut
  $y_{\rm cut}=0$ in the ANN output.
  %
  The $y_{\rm cut}=0$
  values correspond to the  cut-based analysis
  results from Table~\ref{table:cutflow}.
}
\label{fig:sb_mva}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

### 2015-11-17 paper/sec-pileup.tex
Even more work on paper

To conclude, it is illustrative to quantify which of the input variables
to the MVA are most significant in the case of PU,
and compare it with the corresponding
results without PU.
%
We use the same estimator as in Sect.~\ref{sec:signalsignificance},
namely the sum
of the absolute value of all the weights connected to a given
input neuron $i$, Eq.~(\ref{eq:totweight}).
%
The results for the resolved and boosted categories are shown
on Fig. ...
%
As we can see

>>>>

To conclude, it is illustrative to quantify which of the input variables
to the MVA are most significant in the case of PU,
and compare it with the corresponding
results without PU.
%
We use the same estimator as in Sect.~\ref{sec:signalsignificance},
namely the sum
of the absolute value of all the weights connected to a given
input neuron $i$, Eq.~(\ref{eq:totweight}).
%
The results for the resolved and boosted categories are shown
on Fig.~\ref{fig:nnweights_PU}.
%
As we can see by comparing to the corresponding
results without PU in Fig.~\ref{fig:nnweights}, we see that
in the boosted category, the discrimination power of the invariant
mass of Higgs candidates is decreased and that of the various substructure
variables, in particular $C_2$ and $D^{{\rm rm}}$, is conversely
increased.
%
This reflects the fact that these substructure variables are
relatively robust against PU contamination.

In the case of the resolved category, in the case without PU the highest
values of the weight sum were found for the $p_T$ of the leading
Higgs and for the Higgs invariant masses.
%
This is also true in the case with PU, but now the rapidity difference
between the two Higgs candidates, $\Delta y_{hh}$ increases its
importance, again reflecting that this variable is relatively
insensitive to PU effects.

%%%%%%%%%%%%%%%%%%%%%%%%
\begin{figure}[t]
\begin{center}
\includegraphics[width=0.49\textwidth]{plots/res_wgthist_SKPU80.pdf}
\includegraphics[width=0.49\textwidth]{plots/bst_wgthist_SKPU80.pdf}
\vspace{-0.5cm}
\caption{\small
Same as Fig.~\ref{fig:nnweights} in the case of PU.
}
\label{fig:nnweights_PU}
\end{center}
\end{figure}
%%%%%%%%%%%%%%%%%%%%%%%

\subsection{Combined results}

Finally, we now combine the three categories, resolved,
intermediate and boosted, and determine the signal
significance taking into account all event topologies.
%
First of all, in Table~\ref{table:cutflowMVA_PU}
we provide the  the number of signal and
    background events at the HL-LHC, as well as the $S/\sqrt{B}$ and $S/B$,
    for both the cut-based analysis (corresponding
    to $y_{\rm cut}=0$) and after the
    optimal MVA cut.
    %
    These numbers corresponding to the simulations
    with $\la n_{\rm PU}\ra=80$
    and SK subtraction.
    %
    The corresponding results without PU were collected
    in Table.~\ref{table:cutflowMVA}.
    %
    The value of the optimal MVA cut $y_{\rm cut}$ is determined
    as the one leading to highest signal significance, with
    the constraint that number of MC events left is still large
    enough to tame statistical fluctuations.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{table}[t]
  \centering
  \begin{tabular}{c|c|c|c||c|c|c|c}
    \hline
    \multicolumn{8}{c}{Boosted category, $\la n_{\rm PU}\ra=80$+SK}\\
    \hline
    \hline
    \multicolumn{4}{c||}{no MVA cut} & \multicolumn{4}{c}{optimal MVA cut}\\
    \multicolumn{4}{c||}{$y_{\rm cut}=0$} & \multicolumn{4}{c}{$y_{\rm cut}=0.84$}\\
    \hline
    \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$
    & \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$\\
        Signal & Back   &     &   &  Signal & Back   &     &    \\
    \hline
    362  &  $1.4\cdot 10^6$     & 0.31       &  $3\cdot 10^{-4}$ &
    248     &   7250             &  2.9       & 0.034 \\
        \hline
  \end{tabular}
   $\,$\\
  \vspace{0.4cm}
  \noindent
  \begin{tabular}{c|c|c|c||c|c|c|c}
    \hline
    \multicolumn{8}{c}{Intermediate category,  $\la n_{\rm PU}\ra=80$+SK}\\
    \hline
    \hline
    \multicolumn{4}{c||}{no MVA cut} & \multicolumn{4}{c}{optimal MVA cut}\\
    \multicolumn{4}{c||}{$y_{\rm cut}=0$} & \multicolumn{4}{c}{$y_{\rm cut}=0.72$}\\
    \hline
    \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$
    & \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$\\
        Signal & Back   &     &   &  Signal & Back   &     &    \\
    \hline
    146  &    $9\cdot 10^5$   & 0.15        &  $2\cdot 10^{-4}$      &
  53  &  2547        & 1.1        &  0.02 \\
        \hline
  \end{tabular}
    $\,$\\
  \vspace{0.4cm}
  \noindent
  \begin{tabular}{c|c|c|c||c|c|c|c}
    \hline
    \multicolumn{8}{c}{Resolved category,  $\la n_{\rm PU}\ra=80$+SK}\\
    \hline
    \hline
    \multicolumn{4}{c||}{no MVA cut} & \multicolumn{4}{c}{optimal MVA cut}\\
    \multicolumn{4}{c||}{$y_{\rm cut}=0$} & \multicolumn{4}{c}{$y_{\rm cut}=0.49$}\\
    \hline
    \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$
    & \multicolumn{2}{c|}{$N_{\rm ev}$} &  $S/\sqrt{B}$  & $S/B$\\
        Signal & Back   &     &   &  Signal & Back   &     &    \\
    \hline
  1123  &    $8.5\cdot 10^6$   & 0.40       &  $1.3\cdot 10^{-4}$       &
  432  &  $3\cdot 10^4$        &   2.5      & 0.15 \\
        \hline
  \end{tabular}
  \caption{\small The results for the number of signal and
    background events
    at the HL-LHC for both the cut-based analysis and after the
    optimal MVA cut, as well as the $S/\sqrt{B}$ and $S/B$
    values in the two cases, in the case of $\la n_{\rm PU}\ra=80$
    and SK subtraction.
    %
    The corresponding results without PU were collected
    in Fig.~\ref{table:cutflowMVA}.
        \label{table:cutflowMVA_PU}
  }
\end{table}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

From Table~\ref{table:cutflowMVA_PU} we see that, thanks
to the MVA, we can improve the signal significance and the
signal over background ratio from 0.31 and $3\cdot 10^{-4}$
(0.40 and $1.3\cdot 10^{-4}$) up to 2.9 and 0.034 (2.5 and 0.15)
in the boosted (resolved) category.
%
The intermediate category exhibits a lower value of $S/\sqrt{B}$
for the optimal $y_{\rm cut}$ value, around unity.
%
We thus conclude that the boosted category is the most promising
one for the study of Higgs pair production in the $b\bar{b}b\bar{b}$
at the HL-LHC, with however useful information also provided
by the resolved category.

In summary, in this section we have demonstrated the robustness
of our results under the presence of PU conditions as those
expected at the HL-LHC.
